---
title: "Forming a Network from Emails at a Large European Research Institution "
author: "Kelly Ng"
date: "3/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, fig.align = "center")
```

I need the following packages for this project: 
```{r}
library(tidyverse)
library(tidygraph)
library(ggraph)
library(igraph)
```

### 1. Importing data (as tibbles): 
```{r}
eu_nodes <- read_table(
  "email-Eu-core-department-labels.txt.gz",
  col_names = c("person", "dept")
)

eu_edges <- read_table(
  "email-Eu-core.txt.gz",
  col_names = c("from", "to")
)
```

### 2. Cleaning up the data by adding 1 to every person identifier (instead of it being 0).
```{r}
eu_nodes <- mutate(eu_nodes, person = person + 1)

eu_edges <- eu_edges + 1
```

### 3. Creating `tbl_graph` from the data. 
```{r}
eu_netw <- 
  tbl_graph(nodes = eu_nodes, edges = eu_edges) |>
  print()
```

### 4. Initial plot of the complete network.
```{r, cache = TRUE}
ggraph(eu_netw, layout = "kk") +
  geom_edge_bend(
    alpha = 0.1,
    arrow = arrow(length = unit(2, "mm"))
  ) +
  geom_node_point(colour = "brown", alpha = 0.5)
```

### 5. Contracting the network so that in the new network there is: 
- one node for each department 
- one directed edge from department A to department B if A sent an email to B 
- each edge A -> B has an attribute called `weight` that is equal to the number of people in department A who sent an email to department B
```{r}
contracted_eu_igraph <-
  eu_netw |>
  as.igraph() |>
  contract(eu_nodes$dept + 1)
vertex_attr(contracted_eu_igraph, "dept") <-
  0:(gorder(contracted_eu_igraph) - 1)
contracted_directed_netw <-
  as_tbl_graph(contracted_eu_igraph) |>
  convert(to_simple) |>
  activate(edges) |>
  mutate(weight = lengths(.tidygraph_edge_index)) |>
  print()
```

### 6. Making a plot of the contracted network. 
```{r, cache = TRUE}
ggraph(contracted_directed_netw, layout = "stress") +
  geom_edge_fan(
    aes(colour = weight, end_cap = label_rect(node2.dept)),
    alpha = 0.5,
    arrow = arrow(length = unit(0.2, "cm"))
  ) +
  geom_node_label(aes(label = dept), colour = "brown") +
  theme(legend.position = c(1, 1), legend.justification = c(1, 1))
```
This result is not very useful because there are too many edges (lines) connected the nodes that it produces an illegible plot, where readers don't know what is connected to what. However, it is useful knowing the positions of the nodes relative to each other (i.e., 33 node is further away). 

### 7. Visualizing the data as a dendrogram that shows how close relations between departments are. 
```{r}
undirected_simple_netw <- 
  contracted_directed_netw |>
  convert(to_undirected) |>
  convert(to_simple) |>
  activate(edges) |> 
  mutate(weight = map_int(.orig_data, ~ sum(.$weight))) |>
  print()
```

### 8. Using the **tidygraph** function `to_hierarchical_clusters()` to obtain a new netowrk that contains the hierarchy of the dendrogram. 
```{r}
hierarchy <-
  undirected_simple_netw |>
  convert(to_hierarchical_clusters, weights = weight)
hierarchy
```

### 9. Using `ggraph()` to plot a dendogram similar to previously. 
```{r, fig.width = 4, fig.height = 9, out.width = '50%'}
layout <- create_layout(hierarchy, "dendrogram")

ggraph(layout) + 
  geom_edge_elbow() +
  coord_flip() +
  geom_node_text(aes(label = dept, filter = leaf)) + 
  labs(title = "Department hierarchy") 
```

### Evaluating the advantages and disadvantages of representing the network data with the dendogram and network figure. 

**Advantages**

- Can see clearly which node is related to which, unlike the figure in task (6) where too many lines overlapped with each other, which made it illegible.
- It's easy to see what the clusters are present and which nodes are within each cluster. 

**Disadvantages**

- On the dendrogram, we aren't able to know which node is connected to which node, or how many edges each node has as the plot is in a hierarchical structure instead. Hence, the dendrogram needs more explanation. 